---
title: "comparison of RSM and PCM"
author: "Chun Hsien Wu"
date: "2021年12月13日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
2021.12.14

# Partial Credit Model
learning TAM package's function about comparison of RSM and PCM
 website https://bookdown.org/dkatz/Rasch_Biome/polytomous-items.html#polytymous-item-types-and-the-partial-credit-model
```{r}
library(TAM)
setwd("/home/son/刻意練習/統計模擬_實做/Measuring_what_matters")

hls2 <- read.csv("https://raw.githubusercontent.com/danielbkatz/DBER_Rasch-data/master/data/hls_poly_scale.csv")

View(hls2)
```
 hls2's dimension is 317*16
學習用 apply function 把資料集 各個試題答題得分的個數加總 呈現

```{r}
apply(hls2, 2, table)
```
the author mentioned that TAM will automatically run the PCM when our data is polytomous. (Now, I got it!)
```{r,attr.output='style="max-height: 100px;"'}
mod_pcm <- tam(hls2)
summary(mod_pcm)
```

two forms of Partial Credit Model: 

$\theta$ is the latent trait of person $n$.

$X_i$ means person $n$ get $X$ points ($x$ = 0,1, ..., $m_i$) on item i.

$P(X_{ni} = x) = \frac{exp[\sum_{k=0}^x(\theta_n-\delta_{ik})]}{\sum_{h=0}^{m_i}exp[\sum_{k=0}^h(\theta_n-\delta_{ik})]}$

The second form of Partial Credit Model

$P(X_{ni} = x) = \frac{exp[\sum_{k=0}^x(\theta_n-(\delta_{i}+\tau_{ik})]}{\sum_{h=0}^{m_i}exp[\sum_{k=0}^h(\theta_n-(\delta_{i} + \tau_{ik})]}$

$\delta_{i}$是整體試題難度

$\tau_{ik}$ 為試題 $i$ 在第 $j$ 閾值的難度

## PCM (Item difficulty)

這裡 $\delta_{ik}$ 就是 TAM報表中的 xsi
```{r,attr.output='style="max-height: 100px;"'}
# Deltas
xsi <- mod_pcm$xsi
xsi
```
Here, TAM's xsi result is actually $\delta_{ik}$. This is item difficulty.

Note that, $\delta_{ik}=\delta_{i}+\tau_{ik}$. In TAM'result.

beta is exactly $\delta_{i}$, tau.Cat1 is $\tau_{i1}$, tau.Cat2 is $\tau_{i2}$, respectively.

這裡  TAM報表中的beta 就是PCM公式中的 $\delta_i$(整體試題難度)

TAM報表中的 tau.cat1 就是PCM公式中的 $\tau_{i1}$ 第$i$題的第一 階梯難度(step difficulty)或稱為閾難度(threshold difficulty)
```{r,attr.output='style="max-height: 100px;"'}
# Delta-tau parameters
delta_tau <- mod_pcm$item_irt
delta_tau
```

We can see that Hls1_Cat1, Hls1_Cat2 and Hls1_Cat3' value are -1.847, 2.298 and 3.83 respectively.

Now, we can look at Hls1, beta equals 1.4275. It refers to item difficulty. Tau.Cat1, tau.Cat2 and tau.Cat3's value is -3.2736.5, 0.87077 and 2.402828, respectively.

Then Hls1_Cat1 =-1.847 = beta + tau.Cat1 = 1.4275 + -3.273605
```{r}
mod_pcm$item #PCM2 type parameters
```
## PCM (Person ability estimates)
```{r,attr.output='style="max-height: 100px;"'}
WLE.ability.poly <- tam.wle(mod_pcm)
```

```{r}
person.ability.poly <- WLE.ability.poly$theta
head(person.ability.poly)
```

## PCM (Item fit statistics)
```{r}
Fit.poly <- tam.fit(mod_pcm)
```
```{r}
Fit.poly$itemfit
```
## Plot and look at difficulties/thresholds
```{r}
# just look at a few items - items 5-8
plot(mod_pcm, type = "items", export = F, high = 5, items = c(5:8))
```
## extract item steps/adjacent category 
difficulties
```{r}
xsi_pcm <- mod_pcm$xsi
head(xsi_pcm)
```
## Rating Scale 
評定量尺模式的公式如下

$$P(X_{nij} = x) =\frac{exp[\sum_{j=0}^x(\theta_n-\delta_i-\tau_j)]}{\sum_{h=0}^{m_i}exp[\sum_{k=0}^h(\theta_n-\delta_{i} -\tau_j]}$$
若以勝率(odds)對數公式表示則為
$$log\frac{P_{nij}}{P_{ni(j-1)}} = \theta_n - (\delta_i + \tau_j)$$
$\theta_n$表示受試者的能力值;

$\delta_i$為第$i$題的整體難度

$\tau_j$為整份測驗所有試題從得到$j-1$分到$j$分的閾值難度 或稱為階梯難度(step difficulty)
  $P_{nij}$與 $P_{ni(j-1)}$ 表示受試者$n$在第$i$題得到$j$分和$j-1$分的機率

```{r,attr.output='style="max-height: 100px;"'}
# run the Rating Scale Model
mod_rsm <- tam.mml(hls2, irtmodel="RSM")
```

以下可以看到評定量尺模型各試題的難度,同時把部份給分模型模型二的各個試題的整體難度與評定量尺模型各個試題的難度並列加以比較

```{r}
xsi_rsm <- mod_rsm$xsi[1:16,1]
comp_rsm_pcm_diff<-cbind(xsi_rsm, mod_pcm$item[,4])
# add colnames name
colnames(comp_rsm_pcm_diff) <- c("RSM", "PCM")
comp_rsm_pcm_diff
```
## 比較RSM與PCM中 delta-tau值

## 模式比較
