---
title: "comparison of RSM and PCM"
author: "Chun Hsien Wu"
date: "2021年12月13日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
2021.12.14
learning TAM package's function about comparison of RSM and PCM
 website https://bookdown.org/dkatz/Rasch_Biome/polytomous-items.html#polytymous-item-types-and-the-partial-credit-model
```{r}
library(TAM)
setwd("/home/son/刻意練習/統計模擬_實做/Measuring_what_matters")

hls2 <- read.csv("https://raw.githubusercontent.com/danielbkatz/DBER_Rasch-data/master/data/hls_poly_scale.csv")

View(hls2)
```
 hls2's dimension is 317*16
學習用 apply function 把資料集 各個試題答題得分的個數加總 呈現

```{r}
apply(hls2, 2, table)
```
the author mentioned that TAM will automatically run the PCM when our data is polytomous. (Now, I got it!)
```{r}
mod2 <- tam(hls2)
summary(mod2)
```

two forms of Partial Credit Model: $\theta$ is the latent trait of person $n$.

$X_i$ means person $n$ get $X$ points ($x$ = 0,1, ..., $m_i$)on item i

$P(X_{ni} = x) = \frac{exp[\sum_{k=0}^x(\theta_n-\delta_{ik})]}{\sum_{h=0}^{m_i}exp[\sum_{k=0}^h(\theta_n-\delta_{ik})]}$
